\Large\bf FlexibleSUSY --- A spectrum generator generator for supersymmetric models

### 2014-06-09 doc/paper.tex
fix typos

Detailed phenomenological studies have been carried out for scenarios
within the minimal supersymmetric standard model (MSSM).  Such work
has been greatly aided by public spectrum generators for MSSM
\cite{Allanach:2001kg,Porod:2003um,Djouadi:2002ze,Baer:1993ae,Chowdhury:2011zr},
allowing fast and reliable exploration of the sparticle spectrum,
mixings and couplings, which can be obtained from particular choices
of breaking mechanism inspired boundary conditions and specified
parameters. Beyond the MSSM there are also two public spectrum
generators \cite{Ellwanger:2006rn,Allanach:2013kza} for the next to
minimal supersymmetric standard model (NMSSM) \cite{NMSSM} (or for recent
reviews see \cite{Ellwanger:2009dp,Maniatis:2009re}).

@@

These developments motivate the exploration of non-minimal SUSY models
which ameliorate the naturalness problems, as can happen in the \ESSM
\cite{King:2005jy,King:2005my,King:2007uj,Athron:2010zz} and USSM
\cite{Fayet:1977yc,Suematsu:1994qm,Cvetic:1995rj,deCarlos:1997yv,Cvetic:1997ky,Demir:1998dk,Langacker:1998tc,Erler:2002pr,Choi:2006fz,Ham:2007wc,Langacker:2008yv,Ham:2008xf,Kalinowski:2008iq}
by raising the tree level Higgs mass. At the same time they can also
motivates models that are developed with a fresh perspective, based on
other considerations.  In both cases exploration of such models can be
aided if it is possible to quickly create a fast spectrum generator.
Currently there is only one option for this, a SPheno-like FORTRAN
code which can be generated from \sarah
\cite{Staub:2010ty,Staub:2009bi,Staub:2010jh,Staub:2012pb,Staub:2013tta}.

@@

\fs already comes with plenty of predefined models: the CMSSM (simply
called \code{MSSM}), the NMSSM \cite{NMSSM} in it's $Z_3$-symmetric form (called \code{NMSSM}), $Z_3$-violating NMSSM (\code{SMSSM}), the USSM (\code{UMSSM})
\cite{Fayet:1977yc,Suematsu:1994qm,Cvetic:1995rj,deCarlos:1997yv,Cvetic:1997ky,Demir:1998dk,Langacker:1998tc,Erler:2002pr,Choi:2006fz,Ham:2007wc,Langacker:2008yv,Ham:2008xf,Kalinowski:2008iq},
the NUHM-\ESSM (\code{E6SSM}) \cite{Athron:2007en}, the
right-handed neutrino extended MSSM (\code{MSSMRHN}), the NUHM-MSSM
(\code{NUHMSSM}) and the $R$-symmetric MSSM (\code{MRSSM})
\cite{Kribs:2007ac}.  See the content of \code{model_files/} for all
predefined model files.  For all these models spectrum generators can
be generated easily like for the CMSSM in
\secref{sec:quick-start-cmssm}.  The spectrum generator for the
$Z_3$-symmetric NMSSM for example can be generated like this:
%
\begin{lstlisting}[language=bash]
$ ./createmodel --name=NMSSM
$ ./configure --with-models=NMSSM
$ make
\end{lstlisting}%% $
%
One of the design goals is modularity and the possibility to easily
construct custom spectrum generators.  The details of the
customization can be found in Sections
\ref{sec:modfile}--\ref{sec:Flexible}.  As a simple example consider
the NMSSM.  The NMSSM variant above unifies all soft-breaking
trilinear scalar couplings at the GUT scale.  In order to relax this
constraint and use a separate value for $A_\lambda$ at the GUT scale
one can edit the model file \code{model_files/NMSSM/FlexibleSUSY.m.in} and
change the lines
%
\begin{lstlisting}[language=Mathematica]
EXTPAR = { {61, LambdaInput} };

@@

\fs generated NUH \ESSM has also been compared against a handwritten
spectrum generator for a constrained version of the \ESSM
\cite{Athron:2009ue, Athron:2009bs, Athron:2011wu, Athron:2012sq,
  Athron:2012pw, Athron:2013ipa}.  The RGEs could be compared directly
in unit tests and were found to match.  The handwritten code doesn't
include full one loop self energies or tadpoles so tests on these were
not carried.  Although the generators assume different constraints
they could be compared by using the output of the CE$_6$SSM generator
as an input to \fs and the spectra were found to be in reasonable
agreement, given the different levels of precision.

>>>>

Detailed phenomenological studies have been carried out for scenarios
within the minimal supersymmetric standard model (MSSM).  Such work
has been greatly aided by public spectrum generators for the MSSM
\cite{Allanach:2001kg,Porod:2003um,Djouadi:2002ze,Baer:1993ae,Chowdhury:2011zr},
allowing fast and reliable exploration of the sparticle spectrum,
mixings and couplings, which can be obtained from particular choices
of breaking mechanism inspired boundary conditions and specified
parameters. Beyond the MSSM there are also two public spectrum
generators \cite{Ellwanger:2006rn,Allanach:2013kza} for the next to
minimal supersymmetric standard model (NMSSM) \cite{NMSSM} (or for recent
reviews see \cite{Ellwanger:2009dp,Maniatis:2009re}).

@@

These developments motivate the exploration of non-minimal SUSY models
which ameliorate the naturalness problems, as can happen in the \ESSM
\cite{King:2005jy,King:2005my,King:2007uj,Athron:2010zz} and USSM
\cite{Fayet:1977yc,Suematsu:1994qm,Cvetic:1995rj,deCarlos:1997yv,Cvetic:1997ky,Demir:1998dk,Langacker:1998tc,Erler:2002pr,Choi:2006fz,Ham:2007wc,Langacker:2008yv,Ham:2008xf,Kalinowski:2008iq}
by raising the tree level Higgs mass. At the same time they can also
motivate models that are developed with a fresh perspective, based on
other considerations.  In both cases exploration of such models can be
aided if it is possible to quickly create a fast spectrum generator.
Currently there is only one option for this, a SPheno-like FORTRAN
code which can be generated from \sarah
\cite{Staub:2010ty,Staub:2009bi,Staub:2010jh,Staub:2012pb,Staub:2013tta}.

@@

\fs already comes with plenty of predefined models: the CMSSM (simply
called \code{MSSM}), the NMSSM \cite{NMSSM} in it's $Z_3$-symmetric form (called \code{NMSSM}), $Z_3$-violating NMSSM (\code{SMSSM}), the USSM (\code{UMSSM})
\cite{Fayet:1977yc,Suematsu:1994qm,Cvetic:1995rj,deCarlos:1997yv,Cvetic:1997ky,Demir:1998dk,Langacker:1998tc,Erler:2002pr,Choi:2006fz,Ham:2007wc,Langacker:2008yv,Ham:2008xf,Kalinowski:2008iq},
the NUHM \ESSM (\code{E6SSM}) \cite{Athron:2007en}, the
right-handed neutrino extended MSSM (\code{MSSMRHN}), the NUHM-MSSM
(\code{NUHMSSM}) and the $R$-symmetric MSSM (\code{MRSSM})
\cite{Kribs:2007ac}.  See the content of \code{model_files/} for all
predefined model files.  For all these models spectrum generators can
be generated easily like for the CMSSM in
\secref{sec:quick-start-cmssm}.  The spectrum generator for the
$Z_3$-symmetric NMSSM for example can be generated like this:
%
\begin{lstlisting}[language=bash]
$ ./createmodel --name=NMSSM
$ ./configure --with-models=NMSSM
$ make
\end{lstlisting}%% $
%
One of the design goals is modularity and the possibility to easily
construct custom spectrum generators.  The details of the
customization can be found in Sections
\ref{sec:modfile}--\ref{sec:Flexible}.  As a simple example consider
the NMSSM.  The NMSSM variant above unifies all soft-breaking
trilinear scalar couplings at the GUT scale.  In order to relax this
constraint and use a separate value for $A_\lambda$ at the GUT scale
one can edit the model file \code{model_files/NMSSM/FlexibleSUSY.m.in} and
change the lines
%
\begin{lstlisting}[language=Mathematica]
EXTPAR = { {61, LambdaInput} };

@@

\fs generated NUHM \ESSM has also been compared against a handwritten
spectrum generator for a constrained version of the \ESSM
\cite{Athron:2009ue, Athron:2009bs, Athron:2011wu, Athron:2012sq,
  Athron:2012pw, Athron:2013ipa}.  The RGEs could be compared directly
in unit tests and were found to match.  The handwritten code doesn't
include full one-loop self-energies or tadpoles so tests on these were
not carried out.  Although the generators assume different constraints
they could be compared by using the output of the CE$_6$SSM generator
as an input to \fs and the spectra were found to be in reasonable
agreement, given the different levels of precision.
